### 2021-08-06 Main_code/3_Time_series_gen.Rmd
init

---
title: "Input time series generation for the WASA model"
author: "Paolo Colombo"
date: "3/5/2021"
output: html_document
---

# Libraries

```{r libraries}
setwd("C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Analysis")

source("Libraries.R")
source("Functions.R")
```

# Precipitation

## Upload precipitation data

```{r precipitation upload}
setwd("C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Data/meteorological_stations_ceara")

#FUNCEME dataset
input<-list.files(path="./Meteorological_stations_FUNCEME/precipitation_subset",full.names = T,recursive = TRUE)

prec_input<-list()
for(i in 1:length(input)){
  tryCatch(prec_input[[i]]<-read.table(input[i],header = FALSE,
                              sep = "",col.names = c("year","month","day","data"),
                              skip=6),
           error=function(e){cat("ERROR :",conditionMessage(e),i, "\n")})
}
```

## Set up the main dataframe

Creation of the position vector (positions_prec) and the main dataframe containing all the data from the meteorological stations (main_dataframe_prec).

```{r precipitation dataframe}
#Positions vector
for(i in 1:length(input)){
  lines<-readLines(input[i],n = 6)
  ID<-as.numeric(unlist(strsplit(lines[1], ":"))[2])
  name<-unlist(strsplit(lines[2], ":"))[2]
  long<-as.numeric(unlist(strsplit(lines[5], ":"))[2])
  lat<-as.numeric(unlist(strsplit(lines[6], ":"))[2])
  p<-data.frame(ID,name,long,lat)
  if(i > 1) positions_prec<-rbind(positions_prec,p,make.row.names=FALSE)
  else positions_prec<-p
}
positions_prec$ID<-paste0("St",positions_prec$ID)

#Unified dataframe

start_y <- 1980
end_y <- 2020
start_m <- 01
end_m <- 12

start<-paste0("01/0",start_m,"/",start_y)
end<-paste0("31/",end_m,"/",end_y)
date<-seq(as.Date(start,"%d/%m/%Y"), as.Date(end,"%d/%m/%Y"), by="days")

##Create a date column inside each station dataframe
for(i in 1:length(prec_input[])){
  d<-as.Date(paste0(prec_input[[i]]$year,"-",prec_input[[i]]$month,"-",prec_input[[i]]$day))
  prec_input[[i]]$date<-d
}

tic("Unified dataframe - Optimized")
main_dataframe_prec<-data.frame(date)
index<-matrix(NA,ncol = 2,nrow = 744)
j<-1

for(i in 1:length(prec_input[])){
  check<-prec_input[[i]][which(prec_input[[i]]$date %in% main_dataframe_prec$date),]
  if(dim(check)[1]!=0){
      #main_dataframe_prec$new<-rep(NA,nrow(main_dataframe_prec))
      new_v<-sistemadati(vettore = prec_input[[i]],lung = nrow(main_dataframe_prec),g = date)
      main_dataframe_prec$new<-new_v[,2]
      names(main_dataframe_prec)[names(main_dataframe_prec) == "new"]<-positions_prec$ID[i]
  }else{
      index[j,1]<-i
      index[j,2]<-positions_prec$ID[i]
      j<-j+1
  }
}
index_clean<-subset(index, (!is.na(index[,1])) & (!is.na(index[,2])))
toc()
#519.94 sec = 8.65 min
#818.79 sec

remove(index,j,new_v)
View(index_clean) #Stations with no observations

path <- "C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Analysis/time_series_generation/main_datasets"
write.table(main_dataframe_prec,paste0(path,"/main_dataframe_precipitation_3005.txt"),
            quote=FALSE,sep="\t",row.names = FALSE)

```

## Interpolation

```{r precipitation interpolation}
#Upload the main dataframe of precipitation data
# path<-"C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Analysis/time_series_generation/main_datasets"
# main_dataframe_prec<-read.table(paste0(path,"/main_dataframe_precipitation_opt_v2.txt"),
#                                 header = TRUE, sep="\t")

#Import of the subbasins shapefile
path<-"C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Data/shapefile/streamgauges"
subbasins<-readOGR(paste0(path,"/Subbasins_geomfix.shp"))

#Output dataframe creation
tic("Output generation - Precipitation - With function")
output_prec<-create_time_series(main_dataframe = main_dataframe_prec,
                                positions = positions_prec,
                                layer = subbasins,
                                dates = date)
toc()

#7335.77 sec = 2 h
#6608.21 sec
#With new create_time_series function
#
```

```{r check the interpolation procedure}
#Export the generated files to visualize them on QGIS
#Only useful while set upping the interpolation algorithm

shape<-data.frame(t(second),
                      positions_prec$long[which(positions_prec$ID %in% col)],
                      positions_prec$lat[which(positions_prec$ID %in% col)],
                      positions_prec$ID[which(positions_prec$ID %in% col)])
    names(shape)<-c("data","long","lat","ID")

coordinates(shape)<-c("long","lat")
proj4string(shape)<-CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")
shape <- spTransform(shape, raster::crs(subbasins))
path<-"C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Data/meteorological_stations_ceara/Meteorological_stations_FUNCEME/prove_shape"
#raster::shapefile(shape,paste0(path,"/stazioni_temp_conID.shp"))

try<-data_sub
coordinates(try)<-c("x","y")
coordinates(try)<-c("long","lat")
proj4string(try)<-CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")
try <- spTransform(try, raster::crs(subbasins))
path<-"C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Data/meteorological_stations_ceara/Meteorological_stations_FUNCEME/prove_shape"
#raster::shapefile(try,paste0(path,"/subbasins_con_prec_prova2.shp"),overwrite=TRUE)

#raster::shapefile(thiessen,paste0(path,"/thiessen_temp_new.shp"),overwrite=TRUE)

```

### Output dataframe creation and export

```{r precipitation output}
#Insert the date
year<-lubridate::year(main_dataframe_prec$date)
month<-lubridate::month(main_dataframe_prec$date)
month[which(month < 10)]<-paste0("0",month[which(month < 10)])
day<-lubridate::day(main_dataframe_prec$date)
day[which(day < 10)]<-paste0("0",day[which(day < 10)])

output_prec$date<-paste0(day,month,year)

#Insert the number of days
output_prec$`number of days`<-lubridate::yday(main_dataframe_prec$date)

#Round the values
output_exp<-output_prec
output_exp[,3:ncol(output_exp)]<-round(output_exp[,3:ncol(output_exp)], digits = 1)

#Export as an Input file for WASA
path<-"C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Analysis/time_series_generation/output_time_series_gen"

my.write(output_exp,file=paste0(path,"/rain_daily_3005.dat"),
         header = "Daily average precipitation [mmd] for each subasin, ordered according Map-IDs/nDate No. of days Subasin-ID/n0 0 1 2 3 4 5 6 7 8 9 10 11 12 13 14 15 16 17 18 19 20 21 22 23 24 25 26 27 28 29 30 31 32 33 34 35 36 37 38 39 40 41 42 43 44 45 46 47 48 49 50 51 52 53 54 55 56 57 58 59 60 61 62 63 64 65 66 67 68 69 70 71 72 73 74 75 76 77 78 79 80 81 82 83 84 85 86 87 88 89 90 91 92 93 94 95 96 97 98 99 100 101 102 103 104 105 106 107 108 109 110 111 112 113 114 115 116 117 118 119 120 121 122 123 124 125 126 127 128 129 130 131 132 133 134 135 136 137 138 139 140 141 142 143 144 145 146 147 148 149 150 151 152 153 154 155 156 157 158 159 160 161 162 163 164 165 166 167 168 169 170 171 172 173 174 175 176 177 178 179 180 181 182 183 184 185 186 187 188 189 190 191 192 193 194 195 196 197 198 199 200 201 202 203 204 205 206 207 208 209 210 211 212 213 214 215",f = write.table,sepset = FALSE)

#Save it as it is for future use
write.table(output_prec,paste0(path,"/output_prec_save_3005.txt"),quote = FALSE,sep="\t",row.names = FALSE)
```

# Temperature

## Upload temperature data

Since FUNCEME stations present numerous days without temperature data, here INMET stations are also considered. This guarantees at least one observation for each year.

```{r temperature upload}
setwd("C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Data/meteorological_stations_ceara")

#FUNCEME and INMET dataset
input<-list.files(path="./Meteorological_stations_FUNCEME/t2m",full.names = T,recursive = TRUE)

temp_input<-list()
for(i in 1:length(input)){
  tryCatch(temp_input[[i]]<-read.table(input[i],header = FALSE,
                              sep = "",col.names = c("year","month","day","data"),
                              skip=6),
           error=function(e){cat("ERROR :",conditionMessage(e),i, "\n")})
}
```

## Set up the main dataframe

```{r temperature dataframe}
#Positions vector
for(i in 1:length(input)){
  lines<-readLines(input[i],n = 6)
  ID<-as.numeric(unlist(strsplit(lines[1], ":"))[2])
  name<-unlist(strsplit(lines[2], ":"))[2]
  long<-as.numeric(unlist(strsplit(lines[5], ":"))[2])
  lat<-as.numeric(unlist(strsplit(lines[6], ":"))[2])
  p<-data.frame(ID,name,long,lat)
  if(i > 1) positions_temp<-rbind(positions_temp,p,make.row.names=FALSE)
  else positions_temp<-p
}
positions_temp$ID<-paste0("St",positions_temp$ID)

#Unified dataframe

start_y <- 1980
end_y <- 2020
start_m <- 01
end_m <- 12

start<-paste0("01/0",start_m,"/",start_y)
end<-paste0("31/",end_m,"/",end_y)
date<-seq(as.Date(start,"%d/%m/%Y"), as.Date(end,"%d/%m/%Y"), by="days")

##Create a date column inside each station dataframe
for(i in 1:length(temp_input[])){
  d<-as.Date(paste0(temp_input[[i]]$year,"-",temp_input[[i]]$month,"-",temp_input[[i]]$day))
  temp_input[[i]]$date<-d
}

tic("Unified dataframe Temperature - Optimized")

main_dataframe_temp<-data.frame(date)
index<-matrix(NA,ncol = 2,nrow = 744)
j<-1

for(i in 1:length(temp_input[])){
  check<-temp_input[[i]][which(temp_input[[i]]$date %in% main_dataframe_temp$date),]
  if(dim(check)[1]!=0){
      #main_dataframe_temp$new<-rep(NA,nrow(main_dataframe_temp))
      new_v<-sistemadati(vettore = temp_input[[i]],lung = nrow(main_dataframe_temp),g = date)
      main_dataframe_temp$new<-new_v[,2]
      names(main_dataframe_temp)[names(main_dataframe_temp) == "new"]<-positions_temp$ID[i]
  }else{
      index[j,1]<-i
      index[j,2]<-positions_temp$ID[i]
      j<-j+1
  }
}
index_clean<-subset(index, (!is.na(index[,1])) & (!is.na(index[,2])))
toc() #154.5 sec = 2.575 min

remove(index,j,new_v)
View(index_clean)

# path <- "C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Analysis/time_series_generation/main_datasets"
# write.table(main_dataframe_temp,paste0(path,"/main_dataframe_temperature_3005.txt"),
#            quote=FALSE,sep="\t",row.names = FALSE)

```

## Interpolation

```{r temperature interpolation}
# Upload the main dataframe of temperature data
path <- "C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Analysis/time_series_generation/main_datasets"
main_dataframe_temp<-read.table(paste0(path,"/main_dataframe_temperature_3005.txt"),
                               header = TRUE, sep="\t")

#Import of the subbasins shapefile
path<-"C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Data/shapefile/Banabuiu"
subbasins<-readOGR(paste0(path,"/subbasins_cut_geomfix.shp"))

tic("Output generation - Temperature - With function")
output_temp<-create_time_series_Wmean(main_dataframe = main_dataframe_temp,
                                positions = positions_temp,
                                layer = subbasins)
toc()

#Various checks
length(sum(!is.na(main_dataframe_temp[which(lubridate::year(main_dataframe_temp$date)==1986),])))
length(main_dataframe_temp[which(is.na(lubridate::year(main_dataframe_temp$date)==1986)),])

#2610.95 sec = 43 min
#3017.68 sec = 50 min
#2716.2  sec = 45 min
#3682.45 sec
```

### Output dataframe creation and export

```{r temperature output}
#Insert the date
year<-lubridate::year(main_dataframe_temp$date)
month<-lubridate::month(main_dataframe_temp$date)
month[which(month < 10)]<-paste0("0",month[which(month < 10)])
day<-lubridate::day(main_dataframe_temp$date)
day[which(day < 10)]<-paste0("0",day[which(day < 10)])

output_temp$date<-paste0(day,month,year)

#Insert the number of days
output_temp$`number of days`<-lubridate::yday(main_dataframe_temp$date)

#Round the values
output_exp<-output_temp
output_exp[,3:ncol(output_exp)]<-round(output_exp[,3:ncol(output_exp)], digits = 1)

#Export as an Input file for WASA
path<-"C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Analysis/time_series_generation/output_time_series_gen"

my.write(output_exp,file=paste0(path,"/temperature_3005.dat"),
         header = "Daily average temperature (in degree Celcius) for each subasin, ordered according to Map-IDs\nDate No. of days Subasin-ID\n0 0 1 2 3 4 5 6 7 8 9 10 11 12 13 14 15 16 17 18 19 20 21 22 23 24 25 26 27 28 29 30 31 32 33 34 35 36 37 38 39 40 41 42 43 44 45 46 47 48 49 50 51 52 53 54 55 56 57 58 59 60 61 62 63 64 65 66 67 68 69 70 71 72 73 74 75 76 77 78 79 80 81 82 83 84 85 86 87 88 89 90 91 92 93 94 95 96 97 98 99 100 101 102 103 104 105 106 107 108 109 110 111 112 113 114 115 116 117 118 119 120 121 122 123 124 125 126 127 128 129 130 131 132 133 134 135 136 137 138 139 140 141 142 143 144 145 146 147 148 149 150 151 152 153 154 155 156 157 158 159 160 161 162 163 164 165 166 167 168 169 170 171 172 173 174 175 176 177 178 179 180 181 182 183 184 185 186 187 188 189 190 191 192 193 194 195 196 197 198 199 200 201 202 203 204 205 206 207 208 209 210 211 212 213 214 215",f = write.table,sepset = FALSE)

#Save it as it is for future use
write.table(output_temp,paste0(path,"/output_temp_save_3005.txt"),quote = FALSE,sep="\t",row.names = FALSE)

```

# Humidity

## Upload humidity data

Both FUNCEME and INMET dataset are uploaded. Only the stations named "air_humidity" are considered, to avoid duplicated stations.

```{r humidity upload}
setwd("C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Data/meteorological_stations_ceara")

#FUNCEME and INMET dataset
input<-list.files(path="./Meteorological_stations_FUNCEME/humidity_subset",full.names = T,recursive = TRUE)

hum_input<-list()
for(i in 1:length(input)){
  tryCatch(hum_input[[i]]<-read.table(input[i],header = FALSE,
                              sep = "",col.names = c("year","month","day","data"),
                              skip=6),
           error=function(e){cat("ERROR :",conditionMessage(e),i, "\n")})
}
```

## Set up the main dataframe

```{r humidity dataframe}
#Positions vector
for(i in 1:length(input)){
  lines<-readLines(input[i],n = 6)
  ID<-as.numeric(unlist(strsplit(lines[1], ":"))[2])
  name<-unlist(strsplit(lines[2], ":"))[2]
  long<-as.numeric(unlist(strsplit(lines[5], ":"))[2])
  lat<-as.numeric(unlist(strsplit(lines[6], ":"))[2])
  p<-data.frame(ID,name,long,lat)
  if(i > 1) positions_hum<-rbind(positions_hum,p,make.row.names=FALSE)
  else positions_hum<-p
}
positions_hum$ID<-paste0("St",positions_hum$ID)

#Unified dataframe

start_y <- 1980
end_y <- 2020
start_m <- 01
end_m <- 12

start<-paste0("01/0",start_m,"/",start_y)
end<-paste0("31/",end_m,"/",end_y)
date<-seq(as.Date(start,"%d/%m/%Y"), as.Date(end,"%d/%m/%Y"), by="days")

##Create a date column inside each station dataframe
for(i in 1:length(hum_input[])){
  d<-as.Date(paste0(hum_input[[i]]$year,"-",hum_input[[i]]$month,"-",hum_input[[i]]$day))
  hum_input[[i]]$date<-d
}

tic("Unified dataframe humidity - Optimized")

main_dataframe_hum<-data.frame(date)
index<-matrix(NA,ncol = 2,nrow = 744)
j<-1

for(i in 1:length(hum_input[])){
  check<-hum_input[[i]][which(hum_input[[i]]$date %in% main_dataframe_hum$date),]
  if(dim(check)[1]!=0){
      #main_dataframe_hum$new<-rep(NA,nrow(main_dataframe_hum))
      new_v<-sistemadati(vettore = hum_input[[i]],lung = nrow(main_dataframe_hum),g = date)
      main_dataframe_hum$new<-new_v[,2]
      names(main_dataframe_hum)[names(main_dataframe_hum) == "new"]<-positions_hum$ID[i]
  }else{
      index[j,1]<-i
      index[j,2]<-positions_hum$ID[i]
      j<-j+1
  }
}
index_clean<-subset(index, (!is.na(index[,1])) & (!is.na(index[,2])))
toc() #277.06 sec = 4.6 min

remove(index,j,new_v)
View(index_clean)

path <- "C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Analysis/time_series_generation/main_datasets"
write.table(main_dataframe_hum,paste0(path,"/main_dataframe_humidity_8019.txt"),
            quote=FALSE,sep="\t",row.names = FALSE)

```

## Interpolation

```{r humidity interpolation}
#Upload the main dataframe of precipitation data
#path="C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Data/meteorological_stations_ceara/Meteorological_stations_FUNCEME/main_datasets"
#main_dataframe_prec<-read.table(paste0(path,"/main_dataframe_precipitation_opt_v2.txt"),
#                                header = TRUE, sep="\t")

#Import of the subbasins shapefile
path<-"C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Data/shapefile/streamgauges"
subbasins<-readOGR(paste0(path,"/Subbasins_geomfix.shp"))

#Output dataframe creation
tic("Output generation - Humidity - With function")
output_hum<-create_time_series(main_dataframe = main_dataframe_hum,
                                positions = positions_hum,
                                layer = subbasins,
                                dates = date)
toc()

#1437.31 sec = 24 min
```

### Output dataframe creation and export

```{r humidity output}
#Insert the date
year<-lubridate::year(main_dataframe_hum$date)
month<-lubridate::month(main_dataframe_hum$date)
month[which(month < 10)]<-paste0("0",month[which(month < 10)])
day<-lubridate::day(main_dataframe_hum$date)
day[which(day < 10)]<-paste0("0",day[which(day < 10)])

output_hum$date<-paste0(day,month,year)

#Insert the number of days
output_hum$`number of days`<-lubridate::yday(main_dataframe_hum$date)

#Round the values
output_exp<-output_hum
output_exp[,3:ncol(output_exp)]<-round(output_exp[,3:ncol(output_exp)], digits = 1)

#Export as an Input file for WASA
path<-"C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Analysis/time_series_generation/output_time_series_gen"

my.write(output_exp,file=paste0(path,"/humidity_v1.dat"),
         header = "Daily average humidity [in %] for each subasin, ordered according to Map-IDs\nDate No. of days Subasin-ID\n0 0 1 2 3 4 5 6 7 8 9 10 11 12 13 14 15 16 17 18 19 20 21 22 23 24 25 26 27 28 29 30 31 32 33 34 35 36 37 38 39 40 41 42 43 44 45 46 47 48 49 50 51 52 53 54 55 56 57 58 59 60 61 62 63 64 65 66 67 68 69 70 71 72 73 74 75 76 77 78 79 80 81 82 83 84 85 86 87 88 89 90 91 92 93 94 95 96 97 98 99 100 101 102 103 104 105 106 107 108 109 110 111 112 113 114 115 116 117 118 119 120 121 122 123 124 125 126 127 128 129 130 131 132 133 134 135 136 137 138 139 140 141 142 143 144 145 146 147 148 149 150 151 152 153 154 155 156 157 158 159 160 161 162 163 164 165 166 167 168 169 170 171 172 173 174 175 176 177 178 179 180 181 182 183 184 185 186 187 188 189 190 191 192 193 194 195 196 197 198 199 200 201 202 203 204 205 206 207 208 209 210 211 212 213 214 215",f = write.table,sepset = FALSE)

#Save it as it is for future use
path<-"C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Analysis/time_series_generation/output_time_series_gen"
write.table(output_hum,paste0(path,"/output_hum_save_v1.txt"),quote = FALSE,sep="\t",row.names = FALSE)

```

# Radiation

## Upload radiation data

Both FUNCEME and INMET datasets are uploaded

```{r radiation upload}
setwd("C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Data/meteorological_stations_ceara")

#FUNCEME and INMET dataset
input<-list.files(path="./Meteorological_stations_FUNCEME/radiation_KJ_subset/FUNCEME",full.names = T,recursive = TRUE)

rad_input<-list()
for(i in 1:length(input)){
  tryCatch(rad_input[[i]]<-read.table(input[i],header = FALSE,
                              sep = "",col.names = c("year","month","day","data"),
                              skip=6),
           error=function(e){cat("ERROR :",conditionMessage(e),i, "\n")})
}
```

## Set up the main dataframe

```{r radiation dataframe}
#Positions vector
for(i in 1:length(input)){
  lines<-readLines(input[i],n = 6)
  ID<-as.numeric(unlist(strsplit(lines[1], ":"))[2])
  name<-unlist(strsplit(lines[2], ":"))[2]
  long<-as.numeric(unlist(strsplit(lines[5], ":"))[2])
  lat<-as.numeric(unlist(strsplit(lines[6], ":"))[2])
  p<-data.frame(ID,name,long,lat)
  if(i > 1) positions_rad<-rbind(positions_rad,p,make.row.names=FALSE)
  else positions_rad<-p
}
positions_rad$ID<-paste0("St",positions_rad$ID)

#Unified dataframe

start_y <- 1980
end_y <- 2020
start_m <- 01
end_m <- 12

start<-paste0("01/0",start_m,"/",start_y)
end<-paste0("31/",end_m,"/",end_y)
date<-seq(as.Date(start,"%d/%m/%Y"), as.Date(end,"%d/%m/%Y"), by="days")

##Create a date column inside each station dataframe
for(i in 1:length(rad_input[])){
  d<-as.Date(paste0(rad_input[[i]]$year,"-",rad_input[[i]]$month,"-",rad_input[[i]]$day))
  rad_input[[i]]$date<-d
}

tic("Unified dataframe radiation - Optimized")

main_dataframe_rad<-data.frame(date)
index<-matrix(NA,ncol = 2,nrow = 744)
j<-1

for(i in 1:length(rad_input[])){
  check<-rad_input[[i]][which(rad_input[[i]]$date %in% main_dataframe_rad$date),]
  if(dim(check)[1]!=0){
      #main_dataframe_rad$new<-rep(NA,nrow(main_dataframe_rad))
      new_v<-sistemadati(vettore = rad_input[[i]],lung = nrow(main_dataframe_rad),g = date)
      main_dataframe_rad$new<-new_v[,2]
      names(main_dataframe_rad)[names(main_dataframe_rad) == "new"]<-positions_rad$ID[i]
  }else{
      index[j,1]<-i
      index[j,2]<-positions_rad$ID[i]
      j<-j+1
  }
}
index_clean<-subset(index, (!is.na(index[,1])) & (!is.na(index[,2])))
toc() #519.94 sec = 8.65 min

remove(index,j,new_v)
View(index_clean)

path <- "C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Analysis/time_series_generation/main_datasets"
write.table(main_dataframe_rad,paste0(path,"/main_dataframe_radiation_FUNCEME.txt"),
            quote=FALSE,sep="\t",row.names = FALSE)

```

## Interpolation

```{r radiation interpolation}
#Upload the main dataframe of precipitation data
#path="C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Data/meteorological_stations_ceara/Meteorological_stations_FUNCEME/main_datasets"
#main_dataframe_prec<-read.table(paste0(path,"/main_dataframe_radiation_FUNCEME.txt"),
#                                header = TRUE, sep="\t")

#Import of the subbasins shapefile
path<-"C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Data/shapefile/Banabuiu"
subbasins<-readOGR(paste0(path,"/subbasins_cut_geomfix.shp"))

#Output dataframe creation
#Output dataframe creation
tic("Output generation - Radiation - Wmean")
output_rad<-create_time_series_Wmean(main_dataframe = main_dataframe_rad,
                                positions = positions_rad,
                                layer = subbasins)
toc()

#1150.56 sec
#With only Bana reservoirs:
# - Old function: 364.34 sec
# - New function: 586.64 sec
```

### Output dataframe creation and export

ATTENTION
For WASA, the input radiation need to be in W/m2, we have it in KJoule
Make the transformation!
[Here][KJ] they say to multiply the KJ by 3.6. Compare the two time series and see if by multiplying by 3.6 they are in the same range.

[KJ]: https://electronics.stackexchange.com/questions/538921/how-to-convert-calculate-textkj-textm2-to-textw-textm2

```{r scale change}
#Transform from KJ to W/m2

output_rad_Wm2_v1<-output_rad
output_rad_Wm2_v1<-output_rad_Wm2_v1[,3:ncol(output_rad)]*3.6

```

```{r radiation output}
#Insert the date
year<-lubridate::year(main_dataframe_rad$date)
month<-lubridate::month(main_dataframe_rad$date)
month[which(month < 10)]<-paste0("0",month[which(month < 10)])
day<-lubridate::day(main_dataframe_rad$date)
day[which(day < 10)]<-paste0("0",day[which(day < 10)])

output_rad$date<-paste0(day,month,year)

#Insert the number of days
output_rad$`number of days`<-lubridate::yday(main_dataframe_rad$date)

#Round the values
output_exp<-output_rad
output_exp[,3:ncol(output_exp)]<-round(output_exp[,3:ncol(output_exp)], digits = 1)

#Export as an Input file for WASA
path<-"C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Analysis/time_series_generation/output_time_series_gen"

my.write(output_exp,file=paste0(path,"/radiation_v1.dat"),
         header = "Daily average shortwave radiation [in Wm2] for each subasin, ordered according to MAP-IDs\nDate No. of days Subasin-ID\n0 0 1 2 3 4 5 6 7 8 9 10 11 12 13 14 15 16 17 18 19 20 21 22 23 24 25 26 27 28 29 30 31 32 33 34 35 36 37 38 39 40 41 42 43 44 45 46 47 48 49 50 51 52 53 54 55 56 57 58 59 60 61 62 63 64 65 66 67 68 69 70 71 72 73 74 75 76 77 78 79 80 81 82 83 84 85 86 87 88 89 90 91 92 93 94 95 96 97 98 99 100 101 102 103 104 105 106 107 108 109 110 111 112 113 114 115 116 117 118 119 120 121 122 123 124 125 126 127 128 129 130 131 132 133 134 135 136 137 138 139 140 141 142 143 144 145 146 147 148 149 150 151 152 153 154 155 156 157 158 159 160 161 162 163 164 165 166 167 168 169 170 171 172 173 174 175 176 177 178 179 180 181 182 183 184 185 186 187 188 189 190 191 192 193 194 195 196 197 198 199 200 201 202 203 204 205 206 207 208 209 210 211 212 213 214 215",f = write.table,sepset = FALSE)

#Save it as it is for future use
write.table(output_rad,paste0(path,"/output_rad_v1.txt"),quote = FALSE,sep="\t",row.names = FALSE)
```

# Extraterrestrial radiation

Extraterrestrial radiation has a different style compared to the other time series, and it can be obtained easily.

# Comparison between generated and Alexandre's time series

```{r}
setwd("C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Analysis")

source("Libraries.R")
source("Functions.R")

#Load Banabuiu reservoirs' names and IDs
path<-"C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Analysis"
reservoirs<-read.table(paste0(path,"/reservoir_name_ID.txt"), header = TRUE, sep = "\t")
```

## Precipitation visual comparison

```{r}
#Precipitation
path<-"C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Analysis/time_series_generation/output_time_series_gen"
rain_gen<-read.table(paste0(path,"/output_prec_save_v2.txt"),skip = 1)
path<-"C:/Thesis_fortran/WASA/Input/Time_series"
rain_alex<-read.table(paste0(path,"/rain_daily.dat"),skip = 3)
rain_gen<-rain_gen[1:11323,]
rain_alex<-rain_alex[1:11323,] #end in 2010

#Plots
time_index<-seq(as.Date("1980-01-01"),as.Date("2010-12-31"),by="days")
for(i in 1:nrow(reservoirs)){
  #Creation of the dataset for the visualization
  n<-as.numeric(reservoirs$SubbasinID[i])+2
  visual1<-data.frame(rain_alex[,n],rep("Alexandre",length(rain_alex[,n])),time_index)
  names(visual1)<-c("data","id","time")
  visual2<-data.frame(rain_gen[,n],rep("Generated",length(rain_gen[,n])),time_index)
  names(visual2)<-c("data","id","time")
  visual<-cbind(rbind(visual1,visual2))
  Legend<-visual$id
  p<-ggplot(data = visual,
            aes(x = `time`,
                y = `data`))+
    geom_line(aes(colour = Legend),alpha = 0.5) +
    labs(x = "Date",
         y = "Precipitation", 
         title = paste0("Subbasin ",reservoirs$SubbasinID[i]," - ",reservoirs$Reservoir_name[i])) +
    theme_ipsum()
  print(p)
  # path<-"C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Analysis/plots"
  # ggsave(
  #   filename = paste0("0517_prec_",reservoirs$SubbasinID[i],".png"),
  #   plot = p,
  #   device = "png",
  #   path = path,
  #   scale = 1,
  #   width = NA,
  #   height = NA,
  #   units = c("in", "cm", "mm"),
  #   dpi = 300,
  #   limitsize = TRUE
  # )
}
```

## Temperature visual comparison

```{r temperature visual comparison}
#Temperature

#Upload
path<-"C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Analysis/time_series_generation/output_time_series_gen"
temp_gen<-read.table(paste0(path,"/output_temp_save_v3.txt"),skip = 1)
path<-"C:/Thesis_fortran/WASA/Input/Time_series"
temp_alex<-read.table(paste0(path,"/temperature.dat"),skip = 3)
temp_gen<-temp_gen[1:11323,]
temp_alex<-temp_alex[1:11323,] #end in 2010

#Plots
time_index<-seq(as.Date("1980-01-01"),as.Date("2010-12-31"),by="days")
for(i in 1:nrow(reservoirs)){
  #Creation of the dataset for the visualization
  n<-as.numeric(reservoirs$SubbasinID[i])+2
  visual1<-data.frame(temp_alex[,n],rep("Alexandre",length(temp_alex[,n])),time_index)
  names(visual1)<-c("data","id","time")
  visual2<-data.frame(temp_gen[,n],rep("Generated",length(temp_gen[,n])),time_index)
  names(visual2)<-c("data","id","time")
  visual<-cbind(rbind(visual1,visual2))
  Legend<-visual$id
  p<-ggplot(data = visual,
            aes(x = `time`,
                y = `data`))+
    geom_line(aes(colour = Legend),size=0.75,alpha = 0.5) +
    labs(x = "Date",
         y = "Temperature", 
         title = paste0("Subbasin ",reservoirs$SubbasinID[i]," - ",reservoirs$Reservoir_name[i]))+
    theme_ipsum()
  print(p)
  # path<-"C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Analysis/plots"
  # ggsave(
  #   filename = paste0("0517_temp_",reservoirs$SubbasinID[i],".png"),
  #   plot = p,
  #     device = "png",
  #     path = path,
  #     scale = 1,
  #     width = NA,
  #     height = NA,
  #     units = c("in", "cm", "mm"),
  #     dpi = 300,
  #     limitsize = TRUE
  #   )
}
```

## Humidity visual comparison

```{r humidity visual comparison}
#Humidity

#Upload
path<-"C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Analysis/time_series_generation/output_time_series_gen"
hum_gen<-read.table(paste0(path,"/output_hum_save_v1.txt"),skip = 1)
path<-"C:/Thesis_fortran/WASA/Input/Time_series"
hum_alex<-read.table(paste0(path,"/humidity.dat"),skip = 3)
hum_gen<-hum_gen[1:11323,]
hum_alex<-hum_alex[1:11323,] #end in 2010

#Plots
time_index<-seq(as.Date("1980-01-01"),as.Date("2010-12-31"),by="days")

for(i in 1:nrow(reservoirs)){
  #Creation of the dataset for the visualization
  n<-as.numeric(reservoirs$SubbasinID[i])+2
  visual1<-data.frame(hum_alex[,n],rep("Alexandre",length(hum_alex[,n])),time_index)
  names(visual1)<-c("data","id","time")

  visual2<-data.frame(abs(hum_gen[,n]),rep("Generated",length(hum_gen[,n])),time_index)
  names(visual2)<-c("data","id","time")
  visual<-cbind(rbind(visual1,visual2))
  Legend<-visual$id
  p<-ggplot(data = visual,
            aes(x = `time`,
                y = `data`))+
    geom_line(aes(colour = Legend),size=0.75,alpha = 0.5) +
    labs(x = "Date",
         y = "Humidity", 
         title = paste0("Subbasin ",reservoirs$SubbasinID[i]," - ",reservoirs$Reservoir_name[i]))+
    theme_ipsum()
  print(p)
  # path<-"C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Analysis/plots"
  # ggsave(
  #   filename = paste0("0517_hum_",reservoirs$SubbasinID[i],".png"),
  #   plot = p,
  #     device = "png",
  #     path = path,
  #     scale = 1,
  #     width = NA,
  #     height = NA,
  #     units = c("in", "cm", "mm"),
  #     dpi = 300,
  #     limitsize = TRUE
  #   )
}
```

## Radiation visual comparison

```{r radiation visual comparison}
#Radiation

#Upload
path<-"C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Analysis/time_series_generation/output_time_series_gen"
rad_gen<-read.table(paste0(path,"/output_rad_v1.txt"),skip = 1)
path<-"C:/Thesis_fortran/WASA/Input/Time_series"
rad_alex<-read.table(paste0(path,"/radiation.dat"),skip = 3)
rad_gen<-rad_gen[1:11323,]
rad_alex<-rad_alex[1:11323,] #end in 2010

#Chech the coefficient
coeff<-data.frame(reservoirs$SubbasinID)
coeff$mean_coeff<-NA
for(i in 1:nrow(reservoirs)){
  n1<-as.numeric(reservoirs$SubbasinID[i])+2
  n2<-i+2
  div<-rad_alex[,n1]/rad_gen[,n2]
  coeff$mean_coeff[i]<-round(mean(div[is.finite(div)], na.rm = TRUE))
}

#Plots
time_index<-seq(as.Date("1980-01-01"),as.Date("2010-12-31"),by="days")

for(i in 1:nrow(reservoirs)){
  #Creation of the dataset for the visualization
  #n<-as.numeric(reservoirs$SubbasinID[i])+2
  n<-as.numeric(reservoirs$SubbasinID[i])+2
  visual1<-data.frame(rad_alex[,n],rep("Alexandre",length(rad_alex[,n])),time_index)
  names(visual1)<-c("data","id","time")

  n<-i+2
  visual2<-data.frame(abs(rad_gen[,n]),rep("Generated",length(rad_gen[,n])),time_index)
  names(visual2)<-c("data","id","time")
  visual<-cbind(rbind(visual1,visual2))
  Legend<-visual$id
  p<-ggplot(data = visual,
            aes(x = `time`,
                y = `data`))+
    geom_line(aes(colour = Legend),size=0.75,alpha = 0.5) +
    labs(x = "Date",
         y = "Radiation", 
         title = paste0("Subbasin ",reservoirs$SubbasinID[i]," - ",reservoirs$Reservoir_name[i]))+
    theme_ipsum()
  print(p)
  # path<-"C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Analysis/plots"
  # ggsave(
  #   filename = paste0("0517_hum_",reservoirs$SubbasinID[i],".png"),
  #   plot = p,
  #     device = "png",
  #     path = path,
  #     scale = 1,
  #     width = NA,
  #     height = NA,
  #     units = c("in", "cm", "mm"),
  #     dpi = 300,
  #     limitsize = TRUE
  #   )
}
```

```{r}
#Try changing the measuring unit
output_rad_Wm2_v1<-rad_gen
output_rad_Wm2_v1<-output_rad_Wm2_v1[,3:ncol(output_rad)]*3.6

n<-5+2
visual1<-data.frame(rad_alex[,n],rep("Alexandre",length(rad_alex[,n])),time_index)
names(visual1)<-c("data","id","time")
visual2<-data.frame(abs(output_rad_Wm2_v2[,n]),rep("Generated",length(output_rad_Wm2_v2[,n])),time_index)
names(visual2)<-c("data","id","time")
visual<-cbind(rbind(visual1,visual2))
Legend<-visual$id
p<-ggplot(data = visual,
          aes(x = `time`,
              y = `data`))+
    geom_line(aes(colour = Legend),size=0.75,alpha = 0.5) +
    labs(x = "Date",
         y = "Radiation", 
         title = paste0("Subbasin ",reservoirs$SubbasinID[i]," - ",reservoirs$Reservoir_name[i]))+
    theme_ipsum()
print(p)

output_rad_Wm2_v2<-rad_gen
output_rad_Wm2_v2<-output_rad_Wm2_v2[,3:ncol(output_rad_Wm2_v2)]*86.4

```

## Root Mean Square Deviation

RMSE provides a single number that answers the question: "How similar, on average, are the numbers in list1 to list2?".

```{r}

RMSE = function(m, o){
  #m: model values
  #o: observed values
  sqrt(mean((m - o)^2))
}

#Precipitation RMSE
rain_gen[is.na(rain_gen),]<-0

#Define a matrix containing the range of variability for each Sub-basin and its
#RMSE
for(i in 3:ncol(rain_gen)){
  subID<-i-2
  y0<-round(RMSE(rain_gen[,i],rain_alex[,i]), digits = 3)
  y1<-paste0(min(rain_gen[,i])," - ",round(max(rain_gen[,i]),digits = 1))
  y2<-round(mean(rain_gen[,i]), digits = 2)
  y3<-paste0(min(rain_alex[,i])," - ",round(max(rain_alex[,i],digits = 1)))
  y4<-round(mean(rain_alex[,i]), digits = 2)

  if(i > 3){
      prec_comparison<-rbind(prec_comparison, data.frame(subID,y0,y1,y2,y3,y4))
  }else prec_comparison<-data.frame(subID,y0,y1,y2,y3,y4)
}
names(prec_comparison)<-c("SubbasinID","RMSE", "Range generated", "Mean generated", "Range Alexandre", "Mean Alexandre")

#Focus on Banabuiu subbasins

View(prec_comparison[which(prec_comparison$SubbasinID %in% reservoirs$SubbasinID),])
#Save this
path<-"C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Analysis/time_series_generation"
write.table(prec_comparison[which(prec_comparison$SubbasinID %in% reservoirs$SubbasinID),],
            paste0(path,"/RMSE_prec.txt"), row.names = FALSE, sep = "\t")

Bana_prec_RMSE<-read.table(paste0(path,"/RMSE_prec.txt"), header = TRUE, sep = "\t")
View(Bana_prec_RMSE)
```

```{r}
#Temperature RMSE
#Define a matrix containing the range of variability for each Sub-basin and its
#RMSE

temp_gen[is.na(temp_gen)]<-0

for(i in 3:ncol(rain_gen)){
  subID<-i-2
  y0<-round(RMSE(temp_gen[,i],temp_alex[,i]), digits = 3)
  y1<-paste0(min(temp_gen[,i])," - ",round(max(temp_gen[,i]),digits = 1))
  y2<-round(mean(temp_gen[,i]), digits = 2)
  y3<-paste0(min(temp_alex[,i])," - ",round(max(temp_alex[,i],digits = 1)))
  y4<-round(mean(temp_alex[,i]), digits = 2)

  if(i > 3){
      temp_comparison<-rbind(temp_comparison, data.frame(subID,y0,y1,y2,y3,y4))
  }else temp_comparison<-data.frame(subID,y0,y1,y2,y3,y4)
}
names(temp_comparison)<-c("SubbasinID","RMSE", "Range generated", "Mean generated", "Range Alexandre", "Mean Alexandre")

#Focus on Banabuiu subbasins

View(temp_comparison[which(temp_comparison$SubbasinID %in% reservoirs$SubbasinID),])
#Save this
path<-"C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Analysis/time_series_generation"
write.table(temp_comparison[which(temp_comparison$SubbasinID %in% reservoirs$SubbasinID),],
            paste0(path,"/RMSE_temp.txt"), row.names = FALSE, sep = "\t")

Bana_temp_RMSE<-read.table(paste0(path,"/RMSE_temp.txt"), header = TRUE, sep = "\t")

```

# SIGA (abandoned)

* Export the position vector as a point shapefile
* Create a unique dataframe containing all the stations' observations: rows = date, columns = stations. Export it in a .csv file

```{r}
#Export the positions vector as a point shapefile
remove(shp_stations)
shp_stations<-positions_prec[which(positions_prec$ID %in% colnames(main_dataframe_prec)),]
coordinates(shp_stations)<-c("long","lat")

#Import the subbasins shapefile
path<-"C:/Thesis_fortran/WASA/shapefile/streamgauges/Subbasins.shp"
subbasins<-readOGR(path)
summary(subbasins)

#Assign the same reference system as the subbasins shapefile
proj4string(shp_stations)<-CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")
shp_stations <- spTransform(shp_stations, raster::crs(subbasins))
summary(shp_stations)
#EPSG:32724 - WGS 84 / UTM zone 24S - Proiettato

path<-"C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Data/meteorological_stations_ceara/Meteorological_stations_FUNCEME"
raster::shapefile(shp_stations, paste0(path,"/prec_stations_FUNCEME.shp"),overwrite=TRUE)

#Rearrange main_dataframe so SIGA can read it
#1. Insert "-999" where NA are present
#2. Change the date column in DD/MM/AAAA

SIGA_data<-main_dataframe_prec
#insert -999
SIGA_data[is.na(SIGA_data)] <- -999
#change the date column
date_correct<-main_dataframe_prec$date
SIGA_data$date<-format(date_correct,"%d/%m/%Y")

#Save as a .srs file
write.table(SIGA_data,paste0(path,"/SIGA_precipitation.srs"),
            quote=FALSE,sep="\t",row.names = FALSE)

```

# Drafts

## Visualization

```{r}

# Usual area chart
p <- visual %>%
  ggplot( aes(x = time, y = data)) +
  geom_area(aes(colour = Legend), alpha=0.5) +
  geom_line(color="#69b3a2") +
  theme_ipsum()

# Turn it interactive with ggplotly
p <- ggplotly(p)
p

vis<-data.frame(visual1$data,visual2$data)
names(vis)<-c("vis1","vis2")
p <- ggplot(vis, aes(x = x) ) +
  # Top
  geom_density( aes(x = vis1, y = ..density..)) +
  geom_label( aes(x=4.5, y=1, label="variable1")) +
  # Bottom
  geom_density( aes(x = vis2, y = -..density..)) +
  geom_label( aes(x=4.5, y=-1, label="variable2")) +
  theme_ipsum() +
  xlab("value of x")
p

```

```{r old dataset}
#FUNCEME dataset, organized in "Dia 1, Dia 2..."
#input<-list.files(path="./Prec_stations",full.names = T,recursive = TRUE)

#prec_input<-list()
#for(i in 1:length(input)){
#  prec_input[[i]]<-read.table(input[i],header = TRUE,sep = ";")
#}
```

## Clean the data

```{r}
#Remove the stations with non allowed positions (NA or 0)
for(i in 1:length(prec_input[])){
  #If coordinates are 0, remove the station
  if(is.na(prec_input[[i]]$Latitude[1]) || prec_input[[i]]$Latitude[1] == 0){
    prec_input[[i]]<-NULL
  }
  if(i >= length(prec_input[])) break
}

#Replace 888 and 999 with NA
for(i in 1:length(prec_input[])){
    #Put NA instead of 888
    prec_input[[i]]<-na_if(prec_input[[i]],888)
    prec_input[[i]]<-na_if(prec_input[[i]],999)
    #Create a unique ID for each station
    ID<-rep(paste0("Station ",i),length(prec_input[[i]]$Municipios))
    prec_input[[i]]<-data.frame(prec_input[[i]],ID)
}

#Give the stations' ID to the list
for(i in 1:length(prec_input[])){
  names(prec_input)[i]<-prec_input[[i]]$ID[1]
}
```

## Create the matrix and the positions vector

### Positions vector

```{r poition vector}
#Position vector
for(i in 1:length(prec_input[])){
  name<-prec_input[[i]]$Postos[1]
  ID<-paste0("Station ",i)
  lat<-prec_input[[i]]$Latitude[1]
  long<-prec_input[[i]]$Longitude[1]
  if(i !=1 ){
    positions<-rbind(positions, data.frame(name,ID,long,lat))
  }else{
    positions<-data.frame(name,ID,long,lat)
  }
}
```

### Output dataframe creation

```{r}
start_y<-1980; start_m<-1
end_y<-2010;   end_m<-12
test1 <- as.Date("1980-01-01")
test2 <- as.Date("2010-12-31")
test2-test1
ndays<-11322  #result of the difference above

#Import of the subbasins shapefile
path<-"C:/Thesis_fortran/WASA/shapefile/streamgauges/Subbasins.shp"
subbasins<-readOGR(path)

#Output dataframe creation
output<-data.frame(matrix(NA,nrow=ndays,ncol=2+length(subbasins@data$SubbasinID)))
names(output)<-c("date","number of days",t(subbasins@data$SubbasinID))

tic("Output generation")
j<-1
for(year in start_y:end_y){
  for(month in start_m:end_m){
    #Creation of a matrix containing the data for all the stations for 1 month
    st<-0
    for(i in 1:length(prec_input[])){
      v<-prec_input[[i]][which(prec_input[[i]]$Anos == year & prec_input[[i]]$Meses == month),
                         9:ncol(prec_input[[i]])-1]

      if(dim(v)[1] != 0){
        if(st != 0){
          matrix<-data.frame(matrix,t(v))
          st<-append(st,prec_input[[i]]$ID[1])
        }else{
          matrix<-t(v)
          st<-prec_input[[i]]$ID[1]
        }
      }
    }
    names(matrix)<-st
    for(i in 1:nrow(matrix)){
      #First day data, with coordinates associated
      #NA are removed, and stations with NA on that day are not considered
      check<-matrix[i,][which(!is.na(matrix[i,]))]
      row.names(check)<-NULL
      if(ncol(check)>0){
        col<-colnames(check)
        shape<-data.frame(t(check),
                          positions$long[which(positions$ID %in% col)],
                          positions$lat[which(positions$ID %in% col)])
        names(shape)<-c("data","long","lat")

        #Creation of a spatial element
        coordinates(shape)<-c("long","lat")
        crs.pos<-CRS("+proj=longlat") #definition of the projection
        proj4string(shape)<-crs.pos
        shape <- spTransform(shape, crs(subbasins))
        shape@bbox <- subbasins@bbox

        #Interpolate to create a raster
        th <- as(dirichlet(as.ppp(shape)), "SpatialPolygons")
        proj4string(th) <- proj4string(shape)
        th.z <- over(th, shape)
        th.z$data[is.na(th.z$data)] <- 0
        th.spdf <- SpatialPolygonsDataFrame(th, th.z)

        #Superimpose the subbasin shapefile and obtain a mean of the values in the 
        #Thiessen polygons for each subbasin

        data_sub<-over(subbasins,th.spdf,fn=mean)
        #at the spatial locations of object x retrieves the indexes
        #or attributes from spatial object y

        #Save the values in a pre-made data.frame
        output[j,3:ncol(output)]<-t(data_sub)
        j <- j+1
        #}else if(i!=nrow(matrix)){
        #  j <- j+1
      }
    }
    remove(shape,th,th.z,th.spdf,data_sub)
  }
}
toc()
#First try: it took 3h 38m
#Over was done using the function "mean"

```

### Interpolation update

```{r}
#Interpolation update

year<-start_y
month=start_m
i<-1

## Primo tentativo

require(deldir)
require(sp)

del <- deldir(x = shape$long,y = shape$lat,z = shape$data)
del_sp <- tile.list(del)

wcols <- topo.colors(length(del_sp)) 
plot(del_sp, fillcol=wcols, close=TRUE)

#Ok, questo mi crea una tessellazione
#Problema: si può fare analisi spaziale utilizzando questo "oggetto"?

#Prova di soluzione, a questo link: https://stackoverflow.com/questions/9403660/how-to-create-thiessen-polygons-from-points-using-r-packages
polys<-vector(mode='list',length=length(del_sp))
 for (i in seq(along=polys)) {
    pcrds <- cbind(del_sp[[i]]$x, del_sp[[i]]$y)
    pcrds <- rbind(pcrds, pcrds[1,])
    polys[[i]] <- Polygons(list(Polygon(pcrds)), ID=as.character(i))
 }
SP <- SpatialPolygons(polys)
voronoi <- SpatialPolygonsDataFrame(SP,
    data=data.frame(x = shape$long,
    y = shape$lat,z=shape$data))

## Secondo tentativo

voronoipolygons <- function(x) {
  require(deldir)
  require(sp)
  if (.hasSlot(x, 'coords')) {
    crds <- x@coords  
  } else crds <- x
  op <- deldir(x=crds[,2],y= crds[,3],z=crds[,1])
  w <- tile.list(op)
  polys <- vector(mode='list', length=length(w))
  for (i in seq(along=polys)) {
    pcrds <- cbind(w[[i]]$x, w[[i]]$y)
    pcrds <- rbind(pcrds, pcrds[1,])
    polys[[i]] <- Polygons(list(Polygon(pcrds)), ID=as.character(i))
  }
  SP <- SpatialPolygons(polys)
  voronoi <- SpatialPolygonsDataFrame(SP, data=data.frame(x=crds[,2],
    y=crds[,3],z=crds[,1], row.names=sapply(slot(SP, 'polygons'), 
    function(x) slot(x, 'ID'))))
}

tic()
prova<-voronoipolygons(shape)
toc()
#così creo direttamente uno SpatialPolygon
#posso poi applicare il sistema di riferimento ecc, e tagliare
#andare avanti su questa strada!

```

## SIGA first try (not completed)

```{r}
path<-"C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Data/meteorological_stations_ceara"

#Export the positions vector as a point shapefile
shp_stations<-positions
coordinates(shp_stations)<-c("long","lat")
proj4string(shp_stations)<-CRS("+proj=longlat")
raster::shapefile(shp_stations, paste0(path,"/stations.shp"),overwrite=TRUE)
#summary(shp_stations)

#Create a unique dataframe containing all the stations' observations
#Row: date
#Columns: stations

start_y <- 1980
end_y <- 2010
start_m <- 01
end_m <- 12

start<-paste0("01/0",start_m,"/",start_y)
end<-paste0("31/",end_m,"/",end_y)
date<-seq(as.Date(start,"%d/%m/%Y"), as.Date(end,"%d/%m/%Y"), by="days")

main_dataframe<-data.frame(date)
i<-1
for(i in 1:length(prec_input[])){
  for(j in 1:nrow(main_dataframe)){
    y<-lubridate::year(main_dataframe$date[j])
    m<-lubridate::month(main_dataframe$date[j])

    check1<-prec_input[[i]]$Anos[1]
    check2<-prec_input[[i]]$Meses[1]

    if(check1 == y & check2 == m){
      for(a in 1:length(prec_input[[i]])){
        prec_input[[i]][a,8:ncol(prec_input[[i]])]
      }
      break
    }
  }
}

for(month in start_m:end_m){
    #Creation of a matrix containing the data for all the stations for 1 month
    st<-0
    for(i in 1:length(prec_input[])){
      v<-prec_input[[i]][which(prec_input[[i]]$Anos == year & prec_input[[i]]$Meses == month),
                         9:ncol(prec_input[[i]])-1]

      if(dim(v)[1] != 0){
        if(st != 0){
          matrix<-data.frame(matrix,t(v))
          st<-append(st,prec_input[[i]]$ID[1])
        }else{
          matrix<-t(v)
          st<-prec_input[[i]]$ID[1]
        }
      }
    }
    names(matrix)<-st
}
```