### 2022-03-03 Scripts/Drafts/Time_series_processing.Rmd
folders reorganization

---
title: "Time series processing"
author: "Paolo Colombo"
date: "14/6/2021"
output: html_document
---

# Purpose

Processing of the generated time series. At first, remove the outliers, then remove the NAs.

## Functions

```{r functions definition}
check_outlier = function(dataframe, datecol = 1){
  for(i in 1:ncol(dataframe)){
    if(i != datecol){
      Q1<-quantile(dataframe[,i],0.25, na.rm = TRUE)
      Q3<-quantile(dataframe[,i],0.75, na.rm = TRUE)
      IQR<-Q3-Q1
      upper_limit<-Q3 + 1.5*IQR
      lower_limit<-Q1 - 1.5*IQR
      print(i)
      print(sum(dataframe[,i] > upper_limit, na.rm = TRUE)+sum(dataframe[,i] < lower_limit, na.rm = TRUE))
    }
  }
}

remove_outlier = function(dataframe, datecol = 1, fill_val = NA){
  for(i in 1:ncol(dataframe)){
    if(i != datecol){
      Q1<-quantile(dataframe[,i],0.25, na.rm = TRUE)
      Q3<-quantile(dataframe[,i],0.75, na.rm = TRUE)
      IQR<-Q3-Q1
      upper_limit<-Q3 + 1.5*IQR
      lower_limit<-Q1 - 1.5*IQR
      dataframe[,i][dataframe[,i] > upper_limit]<-fill_val
      dataframe[,i][dataframe[,i] < lower_limit]<-fill_val
    }
  }
  dataframe
}

remove_na = function(dataframe, skipcol = 1, k = 7, maxgap = 31){
  #skipcol: number of columns to skip
  firstcol<-1+skipcol
  for(i in firstcol:ncol(dataframe)){
    dataframe[,i]<-imputeTS::na_ma(dataframe[,i], k = k, maxgap = maxgap)
  }
  dataframe
}

```

# Remove outliers

```{r}
#Load the generated time series file

```