### 2022-03-03 Scripts/Core_code/2_DRN_operations/2_New_hdrn.Rmd
folders reorganization

---
title: "Creation of the new HDRN shapefile with capacities associated"
author: "Paolo Colombo"
date: "26/4/2021"
output: html_document
---

# Libraries and functions

```{r libraries}
setwd("C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Analysis")

source("Libraries.R")
source("Functions.R")
```

# Cleaning of the Global Water shapefile

The Global Water was vectorialized from a raster using QGIS. Since it was distributed over two rasters, they were merged and then cut over the Banabuiu area.

## Load of the shapefile

```{r}
setwd("C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Data")

global_water<-readOGR(dsn = "./Global_surface_water/shape/cut_merged.shp")
```

## Remove the big reservoirs from Global Water

```{r}
path<-"C:/Thesis_fortran/WASA/shapefile/Banabuiu"
main_resI<-st_read(paste0(path,"/mainreservoirsI_cut_ref.shp"))
main_resII<-st_read(paste0(path,"/mainreservoirsII_cut_ref.shp"))

threshold<-min(main_resI$AREA)
glob_wat_cut<-global_water[which(global_water$area_geom < threshold),]

```

## Remove DN = 0 from Global Water

DN = 1 means that in that element water is present. Removing the shapes which present DN = 0 is then useful for the next step in QGIS (the NNJoin), so that the HDRN points will not be matched with the areas which don't present water in them.

```{r}
glob_wat_cut<-glob_wat_cut[which(glob_wat_cut$DN == 1),]
```

## Export the "clean" shapefile

```{r}
writeOGR(glob_wat_cut,dsn = "./Global_surface_water/shape", layer = "global_wat_cut",
         driver = "ESRI Shapefile" )
```

# HDRN update

The Global Water shapefile has been joined to the FUNCEME reservoirs shapefile. In the joined output, each reservoir has an associated area. There are reservoirs that are associated to the same water body in the Global Water shapefile, meaning that there wasn't the same number of water bodies in the Global Water shapefile as the number of small reservoirs present in the FUNCEME shapefile. The duplicate with the highest distance from the water body will be assigned to the lowest class of reservoirs, following the hypothesis that if it is not present in the Global Water map, it was too small to be detected.

## Load of the new shapefile

```{r load HDRN to update}
HDRN_update<-readOGR(dsn = "./HDRN/HDRN_update_area_ID.shp")
```

## Calculate capacities and define the reservoir classes

```{r Molle relationship}
# Molle area-volume relationship
#Volume = k*(Area/(a*k))^(a/(a-1))
#Area = a*k*(Volume/k)^((a-1)/a)

k <- damk_Molle <- 1500
a <- alpha_Molle <- 2.7

HDRN_update$capacity<-k*(HDRN_update$join_area_/(a*k))^(a/(a-1)) #m3

#Area = a*k*(100000/k)^((a-1)/a)

#Reservoirs classes definition
# res_class<-c(
#               102179,
#               151360,
#               265735,
#               758010,
#               875797530
#             )
#this classification doesn't make any sense with the values we have
#quantile(HDRN_update$capacity, probs = seq(0, 1, 0.2))

#cap_glob<-k*(glob_wat_cut$area_geom/(a*k))^(a/(a-1))
#quantile(cap_glob, probs = seq(0, 1, 0.2))

#Reservoirs classes definition
#From Mamede et al. 2018
res_class<-c(
              5000, #m3
              25000,
              50000,
              100000,
              max(HDRN_update$capacity)+1
            )

```

## Filtering of the duplicates

Find the HDRN's reservoirs (id) that present the same Global Water reservoir (join_fid). To the one with the highest distance, associate the capacity of the smallest reservoir class

```{r associate lowest res class (m1)}
#First method: associate the lowest class to the points with no associate global water
#1. Find the duplicates
#2. Select the point with the lowest distance
#3. To the points with a distance higher than the minimum, associate the lowest reservoir class

tic("Lowest reservoir class to points with no associate global water")
subset<-HDRN_update[duplicated(HDRN_update$join_fid),]
for(i in 1:length(subset$join_fid)){
  #Take the first duplicate ID
  check1<-subset$join_fid[i] 
  #Take the minimum distance from the points with the same associated ID
  check2<-min(HDRN_update$distance[HDRN_update$join_fid %in% check1])
  #Take the point associated to the minimum distance
  check3<-HDRN_update$id[which(HDRN_update$join_fid %in% check1 & HDRN_update$distance == check2)]

  #Associate the lowest reservoir class to the points with distance higher
  #than check2
  HDRN_update$capacity[which(HDRN_update$join_fid %in% check1
                             & HDRN_update$distance > check2
                             & HDRN_update$id != check3)]<-res_class[1]-0.1
}
remove(subset, check1, check2, check3)
toc()

summary(HDRN_update@data)

#Export the updated HDRN
# writeOGR(HDRN_update,dsn = "./HDRN", layer = "HDRN_afterR_v1",
#          driver = "ESRI Shapefile")
```

```{r remove (m2)}
#Second method: remove the points with no associate global water
#1. Find the duplicates
#2. Select the point with the lowest distance
#3. Remove the points with a distance higher than the minimum

tic("Removal of points with no associated global water")
subset<-HDRN_update[duplicated(HDRN_update$join_fid),] #Finds the duplicates
for(i in 1:length(subset$join_fid)){
  #Take the first duplicate ID
  check1<-subset$join_fid[i] 
  #Take the minimum distance from the points with the same associated ID
  check2<-min(HDRN_update$distance[HDRN_update$join_fid %in% check1])
  #Take the point associated to the minimum distance
  check3<-HDRN_update$id[which(HDRN_update$join_fid %in% check1 & HDRN_update$distance == check2)]

  #Remove the duplicate points with a distance higher than check2
  HDRN_update<-HDRN_update[which(!(HDRN_update$join_fid %in% check1
                                      & HDRN_update$distance > check2
                                      & HDRN_update$id != check3)),]
}
remove(subset, check1, check2, check3)
toc()

summary(HDRN_update@data)

#Export the updated HDRN
# writeOGR(HDRN_update,dsn = "./HDRN", layer = "HDRN_afterR_v2",
#          driver = "ESRI Shapefile" )
```

## Classification

On QGIS, visualized that none of the reservoirs in the highest class are the same as the main reservoirs.

```{r classification}
#Do the classification in the same way as the other script

#Load the updated HDRN
#Method 1 (Unidentified reservoirs associated to the lowest class)
# path<-"C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Data/HDRN/HDRN_afterR_v1.shp"
#Method 2 (Unidentified reservoirs removed)
path<-"C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Data/HDRN/HDRN_afterR_v2.shp"

HDRN_update<-readOGR(path)

#Dataframe creation: first column SubbasinID, others number of reservoirs in the classes
path<-"C:/Thesis_fortran/WASA/shapefile/Banabuiu"
subbasins<-st_read(paste0(path,"/subbasins_cut.shp"))
count_class <- subbasins %>%
                filter(!is.na(SubbasinID)) %>%
                group_by(SubbasinID) %>%
                summarize()
count_class <- data.frame(count_class$SubbasinID)
names(count_class) <- "SubbasinID"

for(i in 1:5){
  if(i != 1){
    c <- HDRN_update@data %>%
          filter(!is.na(SubbasinID)) %>%
          group_by(SubbasinID) %>%
      #   filter(capacity <= res_class[i+1] & capacity > res_class[i]) %>%
      #for Mamede classification: below
          filter(capacity < res_class[i] & capacity >= res_class[i-1]) %>%
          summarize(count=n())
  }else{
    c <- HDRN_update@data %>%
          filter(!is.na(SubbasinID)) %>%
          group_by(SubbasinID) %>%
          #filter(capacity <= res_class[i+1]) %>%
      #for Mamede classification: below
          filter(capacity < res_class[i]) %>%
          summarize(count=n())
  }
  for(a in 1:length(c$SubbasinID)){
    for(b in 1:length(count_class$SubbasinID)){
        if(c$SubbasinID[a]==count_class$SubbasinID[b]) count_class[b,i+1]<-c$count[a]
    }
  }
}
names(count_class)<-c("SubbasinID","class1","class2","class3","class4","class5")
count_class[is.na(count_class)]<-0

#Class ripartition
path<-"C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Analysis/gen_files"
count_class<-read.table(paste0(path, "/lake_number_newHDRN_met2.dat"), skip = 2)

for(i in 2:6){
  print(paste0("Percentage of resevoirs in Class ",i-1))
  print(paste0(round(sum(count_class[,i])/sum(count_class[,2:6]),digits = 3)*100," %"))
}
```

### Creation of lake.dat

```{r lake.dat}
#Create the dataframe
par_class<-data.frame(seq(1,5,1),     #Reservoir_class-ID
                      round(res_class,2),      #maxlake0
                      rep(1,5),       #lake_vol0_factor
                      rep(0,5),       #lake_change: this parameter can be changed in the calibration phase
                      rep(alpha_Molle,5),   
                      rep(damk_Molle,5),
                      c(12.5,18.75,25,37.5,50),       #damc_hrr
                      rep(1.5,5)      #damd_hrr
                      )
names(par_class)<-c("Reservoir_class-ID", "maxlake0_m3", "lake_vol0_factor", "lake_change", "alpha_Molle", "damk_Molle", "damc_hrr", "damd_hrr")
View(par_class)

#Save it
options(scipen = 999)
path<-"C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Analysis/gen_files"
my.write(par_class,file=paste0(path,"/lake.dat"),
         header = "Specification of parameters for the reservoir size classes\nReservoir_class-ID, maxlake0[m**3], lake_vol0_factor[-], lake_change[-], alpha_Molle[-], damk_Molle[-], damc_hrr[-], damd_hrr[-]",f = write.table)
options(scipen = 0)
```

### Creation of lake_number.dat

```{r lake_number}
path<-"C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Analysis/gen_files"
my.write(count_class,file=paste0(path,"/lake_number_newHDRN_met2.dat"),
         header = "Specification of total number of reservoirs in the size classes\nSub-basin-ID, acud[-] (five reservoir size classes)",f = write.table)
```

### Creation of lake_maxvol.dat

For each subbasin, a representative reservoir in each class can be identified, and its starting volume can be provided to WASA in the lake_maxvol.dat file. To identify the representative reservoir for each class, take the median capacity in each reservoir class, for each subbasin.
In the gen_files folder a lake_maxvol computed with the mean can also be found.

```{r representative reservoirs}
path<-"C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Analysis"
res<-read.table(paste0(path,"/reservoir_name_ID.txt"), header = TRUE, sep = "\t")

median_class<- data.frame(res$SubbasinID)
names(median_class) <- "SubbasinID"

for(i in 1:5){
  if(i != 1){
    c <- HDRN_update@data %>%
          filter(!is.na(SubbasinID)) %>%
          group_by(SubbasinID) %>%
      #   filter(capacity <= res_class[i+1] & capacity > res_class[i]) %>%
      #for Mamede classification: below
          filter(capacity < res_class[i] & capacity >= res_class[i-1]) %>%
          summarize(median(capacity))
  }else{
    c <- HDRN_update@data %>%
          filter(!is.na(SubbasinID)) %>%
          group_by(SubbasinID) %>%
          #filter(capacity <= res_class[i+1]) %>%
      #for Mamede classification: below
          filter(capacity < res_class[i]) %>%
          summarize(median(capacity))
  }
  for(a in 1:length(c$SubbasinID)){
    for(b in 1:length(median_class$SubbasinID)){
        if(c$SubbasinID[a]==median_class$SubbasinID[b]) median_class[b,i+1]<-c$`median(capacity)`[a]
    }
  }
}
names(median_class)<-c("SubbasinID","class1","class2","class3","class4","class5")
median_class[is.na(median_class)]<-0
```

```{r export lake_maxvol}
path<-"C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Analysis/gen_files"
my.write(round(median_class,2),file=paste0(path,"/lake_maxvol_newHDRN_m2_median.dat"),
         header = "Specification of water storage capacity for the reservoir size classes\nSub-basin-ID, maxlake[m**3] (five reservoir size classes)",f = write.table)
```

# Scenario without the HDRN

Create lake_number.dat in order to be able to evaluate the scenario without small reservoirs in the WASA model.
Since lake_maxvol.dat is an optional file, it will not be provided to the model when evaluating this scenario.

```{r no HDRN lake}
#Dataframe construction
par_class<-data.frame(seq(1,5,1),     #Reservoir_class-ID
                      round(res_class,2),      #maxlake0
                      rep(1,5),       #lake_vol0_factor
                      rep(0,5),       #lake_change: leave it to 0, so the number of reservoirs doesn't increase
                      rep(alpha_Molle,5),   
                      rep(damk_Molle,5),
                      c(12.5,18.75,25,37.5,50),       #damc_hrr
                      rep(1.5,5)      #damd_hrr
                      )
names(par_class)<-c("Reservoir_class-ID", "maxlake0_m3", "lake_vol0_factor", "lake_change", "alpha_Molle", "damk_Molle", "damc_hrr", "damd_hrr")

#Save it
options(scipen = 999)
path<-"C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Analysis/gen_files"
my.write(par_class,file=paste0(path,"/lake_noHDRN.dat"),
         header = "Specification of parameters for the reservoir size classes\nReservoir_class-ID, maxlake0[m**3], lake_vol0_factor[-], lake_change[-], alpha_Molle[-], damk_Molle[-], damc_hrr[-], damd_hrr[-]",f = write.table)
options(scipen = 0)
```

```{r no HDRN lake_number}
#Sub-basins without networks of small reservoirs must not be entered in the file.
#Create a blank file and try if it works
#Otherwise, create a file with only 1 reservoir for each subbasin, in the lowest class

#No reservoirs
path<-"C:/Users/Utente/OneDrive - Politecnico di Milano/Backup PC/Uni/Thesis/Analysis/gen_files"
write.table("Specification of total number of reservoirs in the size classes\nSub-basin-ID, acud[-] (five reservoir size classes)",paste0(path,"/lake_number_noHDRN_v1.dat"), row.names = FALSE, quote = FALSE, col.names = FALSE)

#Only one resevoir in the smallest reservoir class
count_class[,2]<-1
count_class[,3:6]<-0
my.write(count_class,file=paste0(path,"/lake_number_noHDRN_v2.dat"),
         header = "Specification of total number of reservoirs in the size classes\nSub-basin-ID, acud[-] (five reservoir size classes)",f = write.table)

#Only zeros
#it works!!!!
count_class[,2]<-0
count_class[,3:6]<-0
my.write(count_class,file=paste0(path,"/lake_number_noHDRN_v3.dat"),
         header = "Specification of total number of reservoirs in the size classes\nSub-basin-ID, acud[-] (five reservoir size classes)",f = write.table)

```