### 2022-07-19 R/Figures_code/paper_figures.Rmd
moved data

---
title: Code to generate the figures needed in 'The influence of small reservoirs on
  hydrological drought propagation in space and time'
author: "Paolo Colombo"
date: "08/05/2022"
output:
  html_document: 
    fig_width: 10
    fig_height: 7
  pdf_document: default
editor_options: 
  chunk_output_type: inline
---

# Setup

Set the working directory and load general libraries and custom functions.

```{r, setwd, include = FALSE}
knitr::opts_knit$set(root.dir = "C:/Directory_thesis_codes")
```

```{r, setup, include = FALSE}
setwd("C:/Directory_thesis_codes")
source("./R/Libraries/Libraries.R")
source("./R/Libraries/Libraries_plot.R")
source("./R/Libraries/Functions.R")
source("./R/Libraries/Functions_MC.R")
```

# Figure generation

## Figure 1

Map of the study area localization on the left. Centralized reservoirs visualization and density representation of the DRN on the right. Made in QGIS utilizing the shapefiles and rasters found in ./Data/DNR and ./Data/Shapefile. ![Study area framing. North-East Brazil, Ceará and Banabuiú catchment on the left. On the right, centralized reservoirs location and storage capacity, overlayed to DRN's density in the Banabuiú basin.](C:/Directory_thesis_codes/Figures%20and%20tables/paper/Figure_1.png)

## Figure 2

![Graphical representation of the Drought Cycle Analysis, from Ribeiro Neto et al. (2022). WSI: Water Scarcity Index, PI: Precipitation Index.](C:/Directory_thesis_codes/Figures%20and%20tables/paper/Figure%202.jpg){width="682"}

## Figure 5

```{r figure 5}
setwd("C:/Directory_thesis_codes")

#Custom functions
#Moving average
ma <- function(x, n = 5){stats::filter(x, rep(1 / n, n), sides = 2)}
#Find the lower and upper bounds
find_bounds = function(df, x = 1, skip = 0, ma = FALSE, n = 5){
  minx <- function(x) min(x, na.rm = TRUE)
  maxx <- function(x) max(x, na.rm = TRUE)
  skip <- skip + 1
  lowbound <- apply(df[, skip:ncol(df)], x, minx)
  highbound <- apply(df[, skip:ncol(df)], x, maxx)
  if(ma){
    lowbound <- ma(lowbound, n)
    highbound <- ma(highbound, n)
  }
  bounds <- data.frame(date = df$date, low = lowbound, high = highbound)
  return(bounds)
}

#Cunstruct the dataframe
DCA_AR <- list.load("./Data/Analysis/DCA/DCA_AR.RData")

df <- data.frame(date = DCA_AR$Banabuiu$date)
for(i in seq_len(length(DCA_AR))){
  df$new <- DCA_AR[[i]]$PI
  names(df)[names(df) == "new"] <- names(DCA_AR)[i]
}

newdf <- df[, 1:ncol(df)-1]
bounds <- find_bounds(newdf, 1, 1, ma = TRUE, 6)
bounds[bounds == Inf | bounds == -Inf] <- 0
df <- df[, c(1, ncol(df))]
df$lowbound <- bounds$low
df$highbound <- bounds$high
df$ma <- ma(df$Banabuiu, 6)
df <- df[14:nrow(df), ]

#Plot
col <- c("Mean precipitation" = "#293462",
         "Sub-basin variation" = "#F24C4C")
p <- ggplot() +
  #Add the droughts rectangles
  # geom_rect(data = date_ranges, aes(xmin = from, xmax = to, ymin = -Inf, ymax = Inf), alpha = 0.4) +
  #Add the sub-basins variation
  geom_ribbon(data = df, aes(x = date, ymin = lowbound, ymax = highbound, fill = "Sub-basin variation"), alpha = 0.5) +
  #Add the mean line
  geom_line(data = df, aes(x = date, y = ma, colour = "Mean precipitation"), size = 1.25) +
  #Add the drought indicator lines
  geom_hline(yintercept = 0, color = 'lightblue', linetype = "dashed", size = 1.25, alpha = 0.8) +
  geom_text(aes(as.Date("1985-01-01"), 0, label = "Normal", vjust = -0.5, hjust = 0), size = 6) +
  geom_hline(yintercept = -1, color = 'yellow', linetype = "dashed", size = 1.25, alpha = 0.8) +
  geom_text(aes(as.Date("1985-01-01"), -1, label = "Moderately dry", vjust = -0.5, hjust = 0), size = 6) +
  geom_hline(yintercept = -1.5, color = 'orange', linetype = "dashed", size = 1.25, alpha = 0.8) +
  geom_text(aes(as.Date("1985-01-01"), -1.5, label = "Severely dry", vjust = -0.5, hjust = 0), size = 6) +
  geom_hline(yintercept = -2, color = 'red', linetype = "dashed", size = 1.25, alpha = 0.8) +
  geom_text(aes(as.Date("1985-01-01"), -2, label = "Extremely dry", vjust = -0.5, hjust = 0), size = 6) +
  #Modify the axis
  scale_y_continuous(name = 'SPI-12', limits = c(-4, 4), breaks = seq(-4, 4, 1)) +
  scale_x_date(name = 'Date', date_breaks = "2 years", date_labels = '%Y', limits = as.Date(c("1980-01-01", "2018-12-01")), expand = c(0,0)) +
  #Add the legend
  scale_color_manual(name = "", values = col) +
  scale_fill_manual(name = "", values = col) +
  theme(
        axis.text.x = element_text(size = 16),
        axis.text.y = element_text(size = 20),
        axis.title = element_text(size = 22),
        legend.title = element_text(size = 20),
        legend.text = element_text(size = 18),
        legend.position = "bottom",
        panel.background = element_rect(fill = "white", colour = "grey50")
        )

print(p)

plot.save(p, width = 1920, height = 1080, filename = "./Figures and tables/paper/Figure_5.png")
```

## Figure 6

Volume Deviation visualization for all modeled scenarios. Different relationships between storage conditions and the DNR effect are indicated.

```{r figure 6}
setwd("C:/Directory_thesis_codes")
DCA_AR <- list.load("./Data/Analysis/DCA/DCA_AR.RData")
DCA_LR <- list.load("./Data/Analysis/DCA/DCA_LR.RData")
DCA_SR <- list.load("./Data/Analysis/DCA/DCA_SR.RData")
DCA_N <- list.load("./Data/Analysis/DCA/DCA_N.RData")

#Drought date ranges (retrieved from SPI results)
date_ranges <- data.frame(
  from = as.Date(c("1980-12-01", "1987-05-01", "1992-05-01", "1997-05-01", "2001-02-01", "2005-02-01", "2010-05-01", "2012-05-01", "2014-06-01")),
  to = as.Date(c("1984-03-01", "1988-04-01", "1994-04-01", "1999-12-01", "2002-02-01", "2005-12-01", "2010-12-01", "2014-01-01", "2018-12-01"))
)
#Create the VD dataframe
df <- data.frame(date = DCA_AR[['156']]$date,
                 AR = DCA_AR[['156']]$WSI, LR = DCA_LR[['156']]$WSI,
                 SR = DCA_SR$WSI, N = DCA_N$WSI)
df <- reshape2::melt(df, id.vars = 'date', variable.name = 'Scenario')

p <- ggplot() +
  #Add the drought rectangles
  geom_rect(data = date_ranges, aes(xmin = from, xmax = to, ymin = -Inf, ymax = Inf), alpha = 0.2) +
  #Add the VD lines
  geom_line(data = df, aes(x = date, y = value, linetype = Scenario, col = Scenario), alpha = 0.9, size = 1.25) +
  #VD = 0 line
  geom_hline(yintercept = 0, color = 'tomato2', linetype = "dashed", size = 1, alpha = 0.8) +
  scale_y_continuous(name = 'Volume Deviation (VD)', limits = c(-1, 1)) +
  scale_x_date(name = 'Date', date_breaks = "2 years", date_labels = '%Y', limits = as.Date(c("1980-01-01", "2018-12-01")), expand = c(0,0)) +
  scale_color_paletteer_d("colorBlindness::paletteMartin") +
  theme(axis.text.x = element_text(size = 16),
        axis.text.y = element_text(size = 20),
        axis.title = element_text(size = 22),
        legend.title = element_text(size = 20),
        legend.text = element_text(size = 18),
        legend.position = "bottom")
print(p)

plot.save(p, width = 1920, height = 1080, filename = "./Figures and tables/paper/Figure_6.png")
```

## Figure 8

```{r figure 8a}
setwd("C:/Directory_thesis_codes")
load("./Data/Analysis/Downstreamness/Dsv_AR_new.Rdata")
load("./Data/Analysis/Downstreamness/Dsv_LR_new.Rdata")
load("./Data/Analysis/Downstreamness/Dsv_N.Rdata")
load("./Data/Analysis/Downstreamness/Dsv_SR.Rdata")
df <- data.frame(date = Dsv_AR$date,
                 AR = Dsv_AR$Dsv, LR = Dsv_LR$Dsv, 
                 N = Dsv_N$Dsv, SR = Dsv_SR$Dsv)
df <- reshape2::melt(df, id.vars = 'date', variable.name = 'Scenario')
df$Scenario <- as.character(df$Scenario)

date_ranges <- data.frame(
  from = as.Date(c("1980-12-01", "1987-05-01", "1992-05-01", "1997-05-01", "2001-02-01", "2005-02-01", "2010-05-01", "2012-05-01", "2014-06-01")),
  to = as.Date(c("1984-03-01", "1988-04-01", "1994-04-01", "1999-12-01", "2002-02-01", "2005-12-01", "2010-12-01", "2014-01-01", "2018-12-01"))
)
#Colours
#         AR        LR          N         SR
col <- c("#a651a3", "#429649", "#e2703a", "#046582")
p <- ggplot() +
  geom_rect(data = date_ranges, aes(xmin = from, xmax = to, ymin = -Inf, ymax = Inf), alpha = 0.2) +
  geom_line(data = df, aes(x = date, y = value, colour = Scenario, linetype = Scenario), size = 1.25, alpha = 0.9) +
  scale_y_continuous(name = expression(paste('D'['SV'], " [%]")), limits = c(25, 75)) +
  scale_x_date(name = 'Date', date_breaks = "2 years", date_labels = '%Y', limits = as.Date(c("1980-01-01", "2018-12-01")), expand = c(0,0)) +
  scale_color_manual(values = col) +
  theme(
        axis.text.x = element_text(size = 18),
        axis.text.y = element_text(size = 20),
        axis.title = element_text(size = 22),
        legend.title = element_text(size = 20),
        legend.text = element_text(size = 18),
        legend.position = "bottom"
        )

print(p)

# plot.save(p, width = 1920, height = 1080, filename = "./Figures and tables/paper/Figure_8a.png")
```

```{r figure 8b}
setwd("C:/Directory_thesis_codes")
load("./Data/Analysis/Downstreamness/Dsv_AR_new.Rdata")
load("./Data/Analysis/Downstreamness/Dsv_LR_new.Rdata")
load("./Data/Analysis/Downstreamness/Dsv_N.Rdata")
load("./Data/Analysis/Downstreamness/Dsv_SR_new.Rdata")

df <- data.frame(date = Dsv_AR$date,
                 `AR modified` = Dsv_AR$Dsv_modified, LR = Dsv_LR$Dsv,
                 N = Dsv_N$Dsv, `SR modified` = Dsv_SR$Dsv_modified)
df <- reshape2::melt(df, id.vars = 'date', variable.name = 'Scenario')
df$Scenario <- as.character(df$Scenario)
df$Scenario[df$Scenario == "AR.modified"] <- "AR modified"
df$Scenario[df$Scenario == "SR.modified"] <- "SR modified"

date_ranges <- data.frame(
  from = as.Date(c("1980-12-01", "1987-05-01", "1992-05-01", "1997-05-01", "2001-02-01", "2005-02-01", "2010-05-01", "2012-05-01", "2014-06-01")),
  to = as.Date(c("1984-03-01", "1988-04-01", "1994-04-01", "1999-12-01", "2002-02-01", "2005-12-01", "2010-12-01", "2014-01-01", "2018-12-01"))
)
#Colours
#         AR        LR              N           SR
col <- c("#a651a3", "#429649", "#e2703a", "#046582")
p <- ggplot() +
  geom_rect(data = date_ranges, aes(xmin = from, xmax = to, ymin = -Inf, ymax = Inf), alpha = 0.2) +
  geom_line(data = df, aes(x = date, y = value, colour = Scenario, linetype = Scenario), size = 1.25, alpha = 0.9) +
  scale_y_continuous(name = expression(paste('D'['SV'], " [%]")), limits = c(25, 75)) +
  scale_x_date(name = 'Date', date_breaks = "2 years",
               date_labels = '%Y', limits = as.Date(c("1980-01-01", "2018-12-01")), expand = c(0,0)) +
  scale_color_manual(values = col) +
  theme(
        axis.text.x = element_text(size = 18),
        axis.text.y = element_text(size = 20),
        axis.title = element_text(size = 22),
        legend.title = element_text(size = 20),
        legend.text = element_text(size = 18),
        legend.position = "bottom"
        )

print(p)

plot.save(p, width = 1920, height = 1080, filename = "./Figures and tables/paper/Figure_8b.png")
```