### 2023-08-26 R/Figures_code/paper_figures.Rmd
updates after reviews jhydr

---
title: "Code and information needed to generate the figures needed in 'The influence of small reservoirs on hydrological drought propagation in space and time'"
author: "Paolo Colombo"
date: "08/05/2022"
output:
  html_document: 
    fig_width: 10
    fig_height: 7
  pdf_document: default
editor_options: 
  chunk_output_type: inline
---

# Setup

Set the working directory and load general libraries and custom functions.

```{r, setwd, include = FALSE}
knitr::opts_knit$set(root.dir = "C:/DRN-analysis")
```

```{r, setup, include = FALSE}
setwd("C:/DRN-analysis")
source("./R/Libraries/Libraries.R")
source("./R/Libraries/Libraries_plot.R")
source("./R/Libraries/Functions.R")
source("./R/Libraries/Functions_MC.R")
```

# Figure generation

## Figure 1

Map of the study area localization on the left. Centralized reservoirs visualization and density representation of the DRN on the right. Made in QGIS utilizing the shapefiles and rasters found in ./Data/DNR and ./Data/Shapefile. The density visualization was made with the QGIS “Density Analysis” plugin. The field NUMPOINTs was then divided by the area of each cell ($25 \hspace{0.1cm} km^2$) to obtain the density ($n°small\hspace{0.1cm}reservoirs/km^2$). The equation used in the new field was: "NUMPOINTS"/($area/1000000) ![Study area framing. North-East Brazil, Ceará and Banabuiú catchment on the left. On the right, centralized reservoirs location and storage capacity, overlayed to DRN's density in the Banabuiú basin.](C:/DRN-analysis/Figures and tables/paper/Figure_1.png)

## Figure 2

![Graphical representation of the Drought Cycle Analysis, from Ribeiro Neto et al. (2022). WSI: Water Scarcity Index, PI: Precipitation Index.](C:/DRN-analysis/Figures and tables/paper/Figure_2.jpg){width="682"}

## Figure 4

## Figure 5

SPI-12 visualization. a.) 6-months moving average of the SPI-12 computed on the basin’s average precipitation and variability spectrum across the sub-basins. Spatial distribution of  SPI-12 in Banabuiu watershed sub-basins averaged for b.) 1992-1993 drought and c.) 2012-2018 drought.

```{r figure 5a}
setwd("C:/DRN-analysis")

#Custom functions
#Moving average
ma <- function(x, n = 5){stats::filter(x, rep(1 / n, n), sides = 2)}
#Find the lower and upper bounds
find_bounds = function(df, x = 1, skip = 0, ma = FALSE, n = 5){
  minx <- function(x) min(x, na.rm = TRUE)
  maxx <- function(x) max(x, na.rm = TRUE)
  skip <- skip + 1
  lowbound <- apply(df[, skip:ncol(df)], x, minx)
  highbound <- apply(df[, skip:ncol(df)], x, maxx)
  if(ma){
    lowbound <- ma(lowbound, n)
    highbound <- ma(highbound, n)
  }
  bounds <- data.frame(date = df$date, low = lowbound, high = highbound)
  return(bounds)
}

#Cunstruct the dataframe
DCA_AR <- list.load("./Data/Analysis/DCA/DCA_AR.RData")

df <- data.frame(date = DCA_AR$Banabuiu$date)
for(i in seq_len(length(DCA_AR))){
  df$new <- DCA_AR[[i]]$PI
  names(df)[names(df) == "new"] <- names(DCA_AR)[i]
}

newdf <- df[, 1:ncol(df)-1]
bounds <- find_bounds(newdf, 1, 1, ma = TRUE, 6)
bounds[bounds == Inf | bounds == -Inf] <- 0
df <- df[, c(1, ncol(df))]
df$lowbound <- bounds$low
df$highbound <- bounds$high
df$ma <- ma(df$Banabuiu, 6)
df <- df[14:nrow(df), ]

#Plot
col <- c("Mean precipitation" = "#293462",
         "Sub-basin variation" = "#F24C4C")
p <- ggplot() +
  #Add the droughts rectangles
  # geom_rect(data = date_ranges, aes(xmin = from, xmax = to, ymin = -Inf, ymax = Inf), alpha = 0.4) +
  #Add the sub-basins variation
  geom_ribbon(data = df, aes(x = date, ymin = lowbound, ymax = highbound, fill = "Sub-basin variation"), alpha = 0.5) +
  #Add the mean line
  geom_line(data = df, aes(x = date, y = ma, colour = "Mean precipitation"), size = 1.25) +
  #Add the drought indicator lines
  geom_hline(yintercept = 0, color = 'lightblue', linetype = "dashed", size = 1.25, alpha = 0.8) +
  geom_text(aes(as.Date("1985-01-01"), 0, label = "Normal", vjust = -0.5, hjust = 0), size = 8) +
  geom_hline(yintercept = -1, color = 'yellow', linetype = "dashed", size = 1.25, alpha = 0.8) +
  geom_text(aes(as.Date("1985-01-01"), -1, label = "Moderately dry", vjust = -0.5, hjust = 0), size = 8) +
  geom_hline(yintercept = -1.5, color = 'orange', linetype = "dashed", size = 1.25, alpha = 0.8) +
  geom_text(aes(as.Date("1985-01-01"), -1.5, label = "Severely dry", vjust = -0.5, hjust = 0), size = 8) +
  geom_hline(yintercept = -2, color = 'red', linetype = "dashed", size = 1.25, alpha = 0.8) +
  geom_text(aes(as.Date("1985-01-01"), -2, label = "Extremely dry", vjust = -0.5, hjust = 0), size = 8) +
  #Modify the axis
  scale_y_continuous(name = 'SPI-12', limits = c(-4, 4), breaks = seq(-4, 4, 1)) +
  scale_x_date(name = 'Date', date_breaks = "2 years", date_labels = '%Y', limits = as.Date(c("1980-01-01", "2018-12-01")), expand = c(0,0)) +
  #Add the legend
  scale_color_manual(name = "", values = col) +
  scale_fill_manual(name = "", values = col) +
  theme(
        axis.text.x = element_text(size = 22),
        axis.text.y = element_text(size = 22),
        axis.title.y = element_text(size = 24, face = "bold"),
        axis.title.x = element_blank(),
        legend.title = element_blank(),
        legend.text = element_text(size = 22),
        legend.position = "bottom",
        panel.background = element_rect(fill = "white", colour = "grey50")
        )

print(p)

plot.save(p, width = 1920, height = 1080, filename = "./Figures and tables/paper/Figure_5_new.png")
```

```{r figure 5b}
setwd("C:/Directory_thesis_codes")

sub <- readOGR("./Data/Shapefile/subbasins_cut_geomfix.shp")
SPI_list <- list.load("./Data/Analysis/DCA/SPI_list.RData")
DCA_AR <- list.load("./Data/Analysis/DCA/DCA_AR.RData")

df <- as.data.frame(array(dim = c(length(SPI_list), 3)))
names(df) <- c("SubbasinID", "SPI-1992", "SPI-2012")
df$SubbasinID <- names(SPI_list)

idx1992 <- DCA_AR[["123"]]$date %in% create_date_vector(1992, 1993)
idx2012 <- DCA_AR[["123"]]$date %in% create_date_vector(2012, 2018)
for(name in names(SPI_list)){
  df$`SPI-1992`[df$SubbasinID %in% name] <- mean(SPI_list[[name]][["fitted"]][idx1992], na.rm = T)
  df$`SPI-2012`[df$SubbasinID %in% name] <- mean(SPI_list[[name]][["fitted"]][idx2012], na.rm = T)
}

sub_merge <- merge(sub, df, by = "SubbasinID")

writeOGR(sub_merge, dsn = "./Data/Shapefile/analysis", layer = "subbasin_SPI_all", driver = "ESRI Shapefile")
```

## Figure 6

Volume Deviation visualization for all modeled scenarios. Different relationships between storage conditions and the DNR effect are indicated.

```{r figure 6}
setwd("C:/Directory_thesis_codes")
DCA_AR <- list.load("./Data/Analysis/DCA/DCA_AR.RData")
DCA_LR <- list.load("./Data/Analysis/DCA/DCA_LR.RData")
DCA_SR <- list.load("./Data/Analysis/DCA/DCA_SR.RData")
DCA_N <- list.load("./Data/Analysis/DCA/DCA_N.RData")

#Drought date ranges (retrieved from SPI results)
date_ranges <- data.frame(
  from = as.Date(c("1980-12-01", "1987-05-01", "1992-05-01", "1997-05-01", "2001-02-01", "2005-02-01", "2010-05-01", "2012-05-01", "2014-06-01")),
  to = as.Date(c("1984-03-01", "1988-04-01", "1994-04-01", "1999-12-01", "2002-02-01", "2005-12-01", "2010-12-01", "2014-01-01", "2018-12-01"))
)
#Create the VD dataframe
df <- data.frame(date = DCA_AR[['156']]$date,
                 AR = DCA_AR[['156']]$WSI, LR = DCA_LR[['156']]$WSI,
                 SR = DCA_SR$WSI, N = DCA_N$WSI)
df <- reshape2::melt(df, id.vars = 'date', variable.name = 'Scenario')

p <- ggplot() +
  #Add the drought rectangles
  geom_rect(data = date_ranges, aes(xmin = from, xmax = to, ymin = -Inf, ymax = Inf), alpha = 0.2) +
  #Add the VD lines
  geom_line(data = df, aes(x = date, y = value, linetype = Scenario, col = Scenario), alpha = 0.9, size = 1.25) +
  #VD = 0 line
  geom_hline(yintercept = 0, color = 'tomato2', linetype = "dashed", size = 1, alpha = 0.8) +
  scale_y_continuous(name = 'Volume Deviation (VD)', limits = c(-1, 1)) +
  scale_x_date(name = 'Date', date_breaks = "2 years", date_labels = '%Y', limits = as.Date(c("1980-01-01", "2018-12-01")), expand = c(0,0)) +
  scale_color_paletteer_d("colorBlindness::paletteMartin") +
  theme(axis.text.x = element_text(size = 16),
        axis.text.y = element_text(size = 20),
        axis.title = element_text(size = 22),
        legend.title = element_text(size = 20),
        legend.text = element_text(size = 18),
        legend.position = "bottom")
print(p)

plot.save(p, width = 1920, height = 1080, filename = "./Figures and tables/paper/Figure_6.png")
```

## Figure 7

## Figure 8

Barplot of the drought phases percentages in the four scenarios. Percentage of months in the four drought phases in AR, LR, SR and N scenarios computed for for Arrojado Lisboa.

```{r figure 8}
setwd("C:/Directory_thesis_codes")
source("./R/Libraries/Libraries.R")
source("./R/Libraries/Libraries_plot.R")
source("./R/Libraries/Functions.R")
source("./R/Libraries/Functions_MC.R")

library(scales)

DCA_AR <- list.load("./Data/Analysis/DCA/DCA_AR.RData")
DCA_LR <- list.load("./Data/Analysis/DCA/DCA_LR.RData")
DCA_SR <- list.load("./Data/Analysis/DCA/DCA_SR.RData")
DCA_N <- list.load("./Data/Analysis/DCA/DCA_N.RData")

table <- data.frame(matrix(NA, 5, 4))
names(table) <- c("AR", "LR", "SR", "N")
for(i in seq_len(4)){
  table$AR[i] <- sum(DCA_AR[['156']]$quadrant == i, na.rm = TRUE)
  table$LR[i] <- sum(DCA_LR[['156']]$quadrant == i, na.rm = TRUE)
  table$SR[i] <- sum(DCA_SR$quadrant == i, na.rm = TRUE)
  table$N[i] <- sum(DCA_N$quadrant == i, na.rm = TRUE)
}
table[5, ] <- apply(table, 2, function(x) sum(x, na.rm = TRUE))
table <- apply(table, 2, function(x) x/457*100)
table <- as.data.frame(table)
table <- table[-5,]
table$`Drought phase` <- c(
                  "1. Non-occurrence of drought",
                  "2. Meteorological",
                  "3. Hydro-meteorological",
                  "4. Hydrological")

table <- pivot_longer(table, cols = 1:4, names_to = "Scenario", values_to = "Percentage")

p <- ggplot(table, aes(fill = `Drought phase`, y = Percentage, x = Scenario)) + 
      geom_bar(position = "stack", stat = "identity", 
               alpha = 0.5, colour = "black") +
      scale_fill_manual(values = c("#377EB8", "#FFD92F", "#E41A1C", "#B15928")) +
      scale_y_continuous(breaks = round(seq(0, 100, by = 10))) +
      theme(
          axis.title = element_text(size = 26),
          axis.text.x = element_text(size = 24),
          axis.text.y = element_text(size = 24),
          legend.title = element_text(size = 24, face = "bold"),
          legend.text = element_text(size = 20),
          # legend.position = "bottom",
          panel.background = element_rect(fill = "white"),
          panel.grid.major.x = element_blank(),
          panel.grid.major.y = element_line(colour = alpha("grey50", 0.7))
          )
print(p)

plot.save(p, width = 1920, height = 1440, filename = "./Figures and tables/paper/Figure_8.png")
```

## Figure 9

```{r figure 9a}
setwd("C:/Directory_thesis_codes")
load("./Data/Analysis/Downstreamness/Dsv_AR_new.Rdata")
load("./Data/Analysis/Downstreamness/Dsv_LR_new.Rdata")
load("./Data/Analysis/Downstreamness/Dsv_N.Rdata")
load("./Data/Analysis/Downstreamness/Dsv_SR.Rdata")
df <- data.frame(date = Dsv_AR$date,
                 AR = Dsv_AR$Dsv, LR = Dsv_LR$Dsv, 
                 N = Dsv_N$Dsv, SR = Dsv_SR$Dsv)
df <- reshape2::melt(df, id.vars = 'date', variable.name = 'Scenario')
df$Scenario <- as.character(df$Scenario)

date_ranges <- data.frame(
  from = as.Date(c("1980-12-01", "1987-05-01", "1992-05-01", "1997-05-01", "2001-02-01", "2005-02-01", "2010-05-01", "2012-05-01", "2014-06-01")),
  to = as.Date(c("1984-03-01", "1988-04-01", "1994-04-01", "1999-12-01", "2002-02-01", "2005-12-01", "2010-12-01", "2014-01-01", "2018-12-01"))
)
#Colours
#         AR        LR          N         SR
col <- c("#a651a3", "#429649", "#e2703a", "#046582")
p <- ggplot() +
  geom_rect(data = date_ranges, aes(xmin = from, xmax = to, ymin = -Inf, ymax = Inf), alpha = 0.2) +
  geom_line(data = df, aes(x = date, y = value, colour = Scenario, linetype = Scenario), size = 1.25, alpha = 0.9) +
  scale_y_continuous(name = expression(paste('D'['SV'], " [%]")), limits = c(25, 75)) +
  scale_x_date(name = 'Date', date_breaks = "2 years", date_labels = '%Y', limits = as.Date(c("1980-01-01", "2018-12-01")), expand = c(0,0)) +
  scale_color_manual(values = col) +
  theme(
        axis.text.x = element_text(size = 24),
        axis.text.y = element_text(size = 24),
        axis.title = element_text(size = 24),
        legend.title = element_text(size = 26),
        legend.text = element_text(size = 24),
        legend.position = "bottom"
        )

print(p)

plot.save(p, width = 1920, height = 1080, filename = "./Figures and tables/paper/Figure_8a.png")
```

```{r figure 9b}
setwd("C:/Directory_thesis_codes")
load("./Data/Analysis/Downstreamness/Dsv_AR_new.Rdata")
load("./Data/Analysis/Downstreamness/Dsv_LR_new.Rdata")
load("./Data/Analysis/Downstreamness/Dsv_N.Rdata")
load("./Data/Analysis/Downstreamness/Dsv_SR_new.Rdata")

df <- data.frame(date = Dsv_AR$date,
                 `AR modified` = Dsv_AR$Dsv_modified, LR = Dsv_LR$Dsv,
                 N = Dsv_N$Dsv, `SR modified` = Dsv_SR$Dsv_modified)
df <- reshape2::melt(df, id.vars = 'date', variable.name = 'Scenario')
df$Scenario <- as.character(df$Scenario)
df$Scenario[df$Scenario == "AR.modified"] <- "AR modified"
df$Scenario[df$Scenario == "SR.modified"] <- "SR modified"

date_ranges <- data.frame(
  from = as.Date(c("1980-12-01", "1987-05-01", "1992-05-01", "1997-05-01", "2001-02-01", "2005-02-01", "2010-05-01", "2012-05-01", "2014-06-01")),
  to = as.Date(c("1984-03-01", "1988-04-01", "1994-04-01", "1999-12-01", "2002-02-01", "2005-12-01", "2010-12-01", "2014-01-01", "2018-12-01"))
)
#Colours
#         AR        LR              N           SR
col <- c("#a651a3", "#429649", "#e2703a", "#046582")
p <- ggplot() +
  geom_rect(data = date_ranges, aes(xmin = from, xmax = to, ymin = -Inf, ymax = Inf), alpha = 0.2) +
  geom_line(data = df, aes(x = date, y = value, colour = Scenario, linetype = Scenario), size = 1.25, alpha = 0.9) +
  scale_y_continuous(name = expression(paste('D'['SV'], " [%]")), limits = c(25, 75)) +
  scale_x_date(name = 'Date', date_breaks = "2 years",
               date_labels = '%Y', limits = as.Date(c("1980-01-01", "2018-12-01")), expand = c(0,0)) +
  scale_color_manual(values = col) +
  theme(
        axis.text.x = element_text(size = 24),
        axis.text.y = element_text(size = 24),
        axis.title = element_text(size = 24),
        legend.title = element_text(size = 26),
        legend.text = element_text(size = 24),
        legend.position = "bottom"
        )

print(p)

plot.save(p, width = 1920, height = 1080, filename = "./Figures and tables/paper/Figure_8b.png")
```